SummaryStats

### 2023-04-10 analysis/code/exploration/ExploreOutcomes.Rmd
#4 rework outcome plotting

### Naive plot

ListCitiesNotNA <- function(data, outcome, window=1300:1918){
  ids <- data %>% 
    filter(year %in% window) %>% 
    drop_na({{outcome}}) %>% 
    distinct(city_id) %>% 
    .$city_id
  return(ids)
}

const_cities <- ListCitiesNotNA(data, construction)
allen_cities <- ListCitiesNotNA(data, real_wage)
both <- intersect(const_cities, allen_cities)
colour_vec <- c("gold", "cornflowerblue", "coral")
names(colour_vec) <- allen_cities

PlotOutcomeLines <- function(outcome, colour_vec, dt=data, city_list=NULL, window=1300:1918){
  if (length(city_list) == 0){
    city_list <- ListCitiesNotNA(dt, {{outcome}}, window)
  }

  plot <- dt %>% 
    filter(city_id %in% city_list & year %in% window) %>% 
    mutate(city_id = factor(city_id)) %>% 
    drop_na({{outcome}}) %>% 
    ggplot(aes(x=year, y={{outcome}}))+
    geom_line(aes(colour=city_id))+
    scale_color_manual("City", values=colour_vec)+
    scale_x_continuous(limits=c(min(window), max(window)))

@@

wage_plot <- PlotOutcomeLines(real_wage, colour_vec)
wratio_plot <- PlotOutcomeLines(welfare_ratio, colour_vec)
const_plot <- PlotOutcomeLines(construction, colour_vec, city_list = both)

@@

The line plots work really well for the wage data, but not so much for the 25-yearly construction data. It may look better for the yearly construction data, though.

@@

TopConstCities <- function(data, which=1:8, by="obs", window=1300:1918){
  cities_life <- data %>% 
    filter(year %in% window) %>% 
    group_by(city_id) %>% 
    summarise(
      lifetime_construction = sum(construction, na.rm=TRUE),
      observations = sum(!is.na(construction))
      )

  return(top_ids$city_id[which])
}

Plot4 <- function(which4){
  upper <- 1 + 4*which4
  lower <- 4 + 4*which4
  city_list <- TopConstCities(data, lower:upper)
  colour_vec <- viridis(4)
  names(colour_vec) <- city_list
  PlotOutcomeLines(construction, colour_vec, city_list = city_list)
}

plot1 <- Plot4(0)
plot2 <- Plot4(1)
plot3 <- Plot4(2)
plot4 <- Plot4(3)

plot_grid(plot1, plot2, plot3, plot4)
ggsave("const_four.png", device = "png", path="analysis/output/exploration")

@@

This looks messy as fuck, even for only five cities. Again, I hope that this will look better when the changes happen yearly, not simultaneously every 25 years. Actually, since most cities in most years will have only a single event recorded, I will probably plot it as a histogram and try out various bin sizes. Write a separate Plot function for that.

>>>>

### List cities to plot

```{r}
ListCitiesNotNA <- function(data, outcome, window=1300:1918){
  ids <- data %>% 
    filter(year %in% window) %>% 
    drop_na({{outcome}}) %>% 
    distinct(city_id) %>% 
    .$city_id
  return(ids)
}

ListTopCities <- function(data, which=1:8, by="obs", window=1300:1918){
  cities_life <- data %>% 
    filter(year %in% window) %>% 
    group_by(city_id) %>% 
    summarise(
      lifetime_construction = sum(construction, na.rm=TRUE),
      observations = sum(!is.na(construction))
      )

  return(top_ids$city_id[which])
}
```

```{r}
# city list for comparison
const_cities <- ListCitiesNotNA(data, construction)
allen_cities <- ListCitiesNotNA(data, real_wage)
both <- intersect(const_cities, allen_cities)

# city lists for exploring construction
set.seed(151101)
top9 <- ListTopCities(data, which=1:9)
top16 <- ListTopCities(data, which=1:16)
random9 <- sample(const_cities, 9)
random16 <- sample(const_cities, 16)

### Plotting functions

```{r}
PlotOutcomeLines <- function(data, citylist=NULL, outcome, colour_vec, window=1300:1918){
  if (length(citylist) == 0){
    city_list <- ListCitiesNotNA(data, {{outcome}}, window)
  }
  plot <- data %>% 
    filter(city_id %in% citylist & year %in% window) %>% 
    mutate(city_id = factor(city_id)) %>% 
    drop_na({{outcome}}) %>% 
    ggplot(aes(x=year, y={{outcome}}))+
    geom_line(aes(colour=city_id))+
    scale_color_manual("City", values=colour_vec)+
    scale_x_continuous(limits=c(min(window), max(window)))

@@

PlotHistograms <- function(data, citylist, binw=10){
  data |>
    filter(city_id %in% citylist) |> 
    mutate(city_id = as_factor(city_id)) |> 
    filter(construction > 0) |> 
    ggplot(aes(x=year))+
    geom_histogram(aes(y=after_stat(density)), binwidth = binw)+
    geom_density(aes(y=after_stat(density), colour="red"), show.legend = FALSE)+
    facet_wrap(~city_id)
}

PlotDensity <- function(data, citylist, colour_vec, window=1300:1918){
  data |>
    filter(city_id %in% citylist & year %in% window) |> 
    mutate(city_id = as_factor(city_id)) |> 
    filter(construction > 0) |> 
    ggplot(aes(x=year))+
    geom_density(aes(y=after_stat(density), colour=city_id))+
    scale_color_manual("City", values=colour_vec)+
    scale_x_continuous(limits=c(min(window), max(window)))
}

### Comparison to real wages

```{r plots for Allen cities}
colour_vec <- c("gold", "cornflowerblue", "coral")
names(colour_vec) <- allen_cities

wage_plot <- PlotOutcomeLines(data, citylist=both, real_wage, colour_vec, window=1400:1800)
wratio_plot <- PlotOutcomeLines(data, citylist=both, welfare_ratio, colour_vec, window=1400:1800)
const_plot <- PlotDensity(data, citylist = both, colour_vec, window=1400:1800)

@@

@@

PlotHistograms(data, random16) +
  ggtitle("Construction activity: random cities")
PlotHistograms(data, top16) +
  ggtitle("Construction activity: cities with the most observations")

@@

I want to see how the number of recorded events is related to
- a city's population in 1875
- a city's age
